Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriterial problems

### 2024-06-29 2024/OPTIMA/OPTIMA_2024_MCO.tex
OPTIMA fixed bugs

In each series of experiments for these optimization algorithms, the coefficients $\alpha$ from (\ref{eq:16}) changed in the range $[0.1, 0.9]$. Comparison with the algorithm MGSA that does not use machine learning methods (in the characteristic ($\alpha$) $\alpha = 0$) and with the algorithm ML\_MGSA\_Dist taking into account the distance to the separating hyperplane was carried out. The comparison was made on the base of  the calculated HV index characterizing quality of the obtained Pareto set approximation \cite{ML_MCO_2023,Evtushenko2014}. Further subsections include the results of only those calculations in which the $\alpha$ parameter provided the best quality of the solution found. Tables contain results of solving all the problems (\ref{eq:07}) with fixed accuracy $\varepsilon=0.01$ in the termination condition (\ref{eq:14}). Figures present results for a series of decreasing $\varepsilon$ which generates increasing numbers of trials.

@@

\multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}ML\_MGSA\_Prob \\ (Fixed weights)\end{tabular}} & Linear                & 0.04                           & 114.72                 & 963.05              & 2.36                   \\

@@

\multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}ML\_MGSA\_Prob \\ (Adj. weights)\end{tabular}}  & Linear                & 0.04                           & 114.76                 & 1018.03             & 2.23                   \\

@@

\multirow{3}{*}{\begin{tabular}[c]{@{}c@{}}ML\_MGSA\_Prob \\ (Log Norm)\end{tabular}}      & Linear                & 0.05                           & 114.72                 & 711.44              & 3.19                   \\

@@

Compared to MGSA, the other algorithms reduce significantly the number of iterations in conditions when HV-indices are similar. However, the iteration number when using the models with probabilities (see last 3 rows of Table \ref{tab:1} at section ``Experiments for the case of fixed weights'') is greater than for the algorithm based on the separating hyperplane.

>>>>

In each series of experiments for these optimization algorithms, the coefficients $\alpha$ from (\ref{eq:16}) changed in the range $[0.1, 0.9]$. Comparison with the algorithm MGSA that does not use machine learning methods (in the characteristic (\ref{eq:16}) $\alpha = 0$) and with the algorithm ML\_MGSA\_Dist taking into account the distance to the separating hyperplane was carried out. The comparison was made on the base of  the calculated HV index characterizing quality of the obtained Pareto set approximation \cite{ML_MCO_2023,Evtushenko2014}. Further subsections include the results of only those calculations in which the $\alpha$ parameter provided the best quality of the solution found. Tables contain results of solving all the problems (\ref{eq:07}) with fixed accuracy $\varepsilon=0.01$ in the termination condition (\ref{eq:14}). Figures present results for a series of decreasing $\varepsilon$ which generates increasing numbers of trials.

@@

\multirow{3}{*}{ML\_MGSA\_Prob (Fixed weights)} & Linear                & 0.04                           & 114.72                 & 963.05              & 2.36                   \\

@@

\multirow{3}{*}{ML\_MGSA\_Prob (Adj. weights)}  & Linear                & 0.04                           & 114.76                 & 1018.03             & 2.23                   \\

@@

\multirow{3}{*}{ML\_MGSA\_Prob (Log Norm)}      & Linear                & 0.05                           & 114.72                 & 711.44              & 3.19                   \\

@@

Compared to MGSA, the other algorithms reduce significantly the number of iterations in conditions when HV-indices are similar. However, the iteration number when using the models with probabilities (see rows ``ML\_MGSA\_Prob (Fixed weights)'' of Table \ref{tab:1}) is greater than for the algorithm based on the separating hyperplane.
